\chapter{Results}

\section{Evaluation}

\section{Architecture}

\section{Performance}

\section{Precision and Symbolic Computations}

\chapter{Future Work}\label{chap:fw}

Our project is far from a fully-fledged library, and many more modules must be added.

First of all we do not have support for the floating rate bonds
(see appendix X section something) due to the lack of of Monte Carlo
simulation module. A natural next step after floating rate support
would then be to implement interest rate derivatives such as swaps.

There are two obvious ways in which we can extend HQL with interest rate
derivatives support. For instance, swaps can be modelled by combining a
fixed and a floating leg, the latter still not yet being supported. On the
other hand, it could be built as a list of forward-rate agreements (make sure
this is explained above, or refer to Appendix) between the two parties in
question.\\

Alas, our library is also missing a key component which is a proper calendar
module. We currently allow the users to supply a variable number of
settlements per year. This needs to be constrained so that the users either
specify a standard settlement pattern (missing an example) or \emph{all}
payment dates for OTC products. Moreover, HQL currently assumes that all
weekdays are valid payments dates. This doesn't cater to local holidays
(e.g. January $1^{\text{st}}$ in Denmark, $4^{\text{th}}$ of July in the
United States, etc) which may be computed as offsets from Easter Sunday.\\

Another interesting extension is to implement a way of generating term
sructures from portfolios using bootstrapping and interpolation\footnote{need
a reference to Björk/Munk here}. This would essentially allow users to import
data from external sources and use HQL to build term structures and perform
bond valuation.\\

Support for mortgage-backed bonds is also missing in our library. We provide
an MBO class to be used in the future.\\
%Prepayment estimates must also be computed.

%Cubic splines or Nelson-Siegel method for obtain term structure as a function.

Finally, it would be desirable to implement an embedded domain-specific language
for construction of over-the-counter (OTC) products.
As OTC products are subject to counterparty risk, this would be another topic
to investigate.

%The class hierarchichy should be extended to define interfaces for derivatives,
%commodities, money markets etc.. Draw diagram?

\chapter{Conclusion}
